Exponential Integrators for PDEs arising in meteorology

### 2024-09-23 article_4.tex
Mass Matrix and finite elements

Below we compare the time required for the different algorithms to get below an error of $10^6$ and $10^{10}$ when computing $e^Av$ for various matrix sizes.
\begin{figure}[H]
\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-06.png}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-10.png}
\end{figure}
Again we continue to see that the Lanzcos algorithm outperforms the Arnodli algorithm.
\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{References.bib}

>>>>

Below we compare the time required for the different algorithms to get below an error of $10^6$ and $10^{10}$ when computing $e^Av$ for various matrix sizes.
\begin{figure}[H]
\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-06.png}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-10.png}
\end{figure}
Again we continue to see that the Lanzcos algorithm outperforms the Arnodli algorithm.

\section{Application to Finite Element Methods}
Here we discuss the application of the exponential integrator to finite element methods.

We begin with a finite element problem in the weak form:
\begin{align*}
\int \frac d{dt}u_j v_j - R(u)_j v_j dx &= \mathcal{L}(u,v)\\
\text{substituting } u(t,x) &= \sum_{i=1}^n u_i(t)v_i(x)\\
\text{we get } \int\sum_{i=1}^n(\frac d{dt}u_i(t)v_i(x)v_j(x))dx - R_jv_j(x) &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
\frac d{dt}u_i(t)\sum_{i=1}^n\int(v_i(x)v_j(x))dx - R_jv_j(x) &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
M\frac d{dt}u(t) - R &= Lu\\
\frac d{dt}u(t) &= M^{-1}(Lu + R(u))
\end{align*}
A clear limiting factor when computing $M^{-1}(Lu + R(u))$ is taking the inverse of $M$, the matrix.
One way around this is to perform mass lumping on the matrix, where we sum up the values along each row and place them along the diagonal.
From here computing the inverse is trivial.

\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{References.bib}
